The Design and Implementation of the Synchronous Language \CEU

### 2016-10-01 paper.tex
(>) static checks algorithm (>) flat structure

\CEU aims to offer a simpler semantics than Esterel with more determinism and 
fine-grained control over program execution.
%
The following list summarizes the contributions of our design in this 
direction:
%
\begin{itemize}
%
\item \emph{Event-triggered notion of time bound to the semantics of the 
language.}
Event-driven programming is popular in many domains, such as server and GUI 
development.
We believe that programmers are more familiar with dealing with events in 
isolation, which simplifies the reasoning about concurrency.
In addition, the uniqueness of external events is a prerequisite for the 
concurrency checks of \CEU.
%
\item \emph{Deterministic intra-reaction execution and communication.}
Determinism in \CEU does not depend on additional levels of static analysis.
It encompasses the whole language, including memory accesses, system calls, and 
stack-based internal events.
Programmers can always figure out which statement executes next, making runtime 
analysis and debugging easier.
%
\item \emph{Static concurrency checks.}
Although execution is deterministic, the \CEU compiler still advises about 
suspicious statements that can react concurrently to the same event. 
%appropriate serialization.
%
\item \emph{Safe integration with \C.}
When dealing with concurrent system calls, programmers can provide annotations 
to reduce false positives in the static checks, or to force non commutative 
concurrent behavior.
\CEU also requires finalization clauses to handle pointers representing 
resources.
%
\item \emph{First-class timers with dedicated syntax and automatic 
synchronization.}
Given the omnipresence of timers in embedded systems, a dedicated syntax can 
simplify the development and readability of programs.
Furthermore, automatic synchronization releases the programmer from the burden 
of adjusting timers in sequence and in parallel.
%
\end{itemize}

@@

The scheduler executes in two passes:
%
in the \emph{broadcast} pass, it sets all trails that are waiting for the 
current event to the state \code{STACKED} in the current numeric stack level;
%
in the \emph{dispatch} pass, it executes each trail that is \code{STACKED} to 
run in the current level, setting it immediately to \code{INACTIVE}.
%(the trail segment may reset it in sequence if it doesn't terminate).

>>>>

\CEU aims to offer a simpler semantics than Esterel with more determinism and 
fine-grained control over program execution.
%
The following list summarizes the contributions of our design in this 
direction:
%
\begin{itemize}
%
\item \emph{Event-triggered notion of time bound to the semantics of the 
language.}
Event-driven programming is popular in many domains, such as server and GUI 
development.
We believe that programmers are more familiar with dealing with events in 
isolation, which simplifies the reasoning about concurrency.
In addition, the uniqueness of external events is a prerequisite for the 
concurrency checks of \CEU.
%
\item \emph{Deterministic intra-reaction execution and communication.}
Determinism in \CEU does not depend on additional levels of static analysis.
It encompasses the whole language, including memory accesses, system calls, and 
stack-based internal events.
Programmers can figure out which statement executes next, helping runtime 
analysis and debugging.
%
\item \emph{Static concurrency checks.}
Although execution is deterministic, the \CEU compiler still advises about 
suspicious statements that can react concurrently to the same event. 
%appropriate serialization.
%
\item \emph{Safe integration with \C.}
When dealing with concurrent system calls, programmers can provide annotations 
to reduce false positives in the static checks, or to force non commutative 
concurrent behavior.
\CEU also requires finalization clauses to handle pointers representing 
resources.
%
\item \emph{First-class timers with dedicated syntax and automatic 
synchronization.}
Given the omnipresence of timers in embedded systems, a dedicated syntax can 
simplify the development and readability of programs.
Furthermore, automatic synchronization releases the programmer from the burden 
of adjusting timers in sequence and in parallel.
%
\end{itemize}

@@

The scheduler executes in two passes:
%
in the \emph{broadcast} pass, it sets all trails that are waiting for the 
current event to the state \code{STACKED} in the current numeric stack level;
%
in the \emph{dispatch} pass, it executes each trail that is \code{STACKED} to 
run in the current level, setting it immediately to \code{INACTIVE}.
%(the trail segment may reset it in sequence if it doesn't terminate).
%
Since the trail space is flat and carries no information about the program
hierarchy (ln. \bx:12--15), all trails (active or inactive) are traversed in
all passes.
We believe that this makes reaction times more predictable and improves
worst-case reaction times (e.g., with all trails active).
However, we did not investigate these assumptions.
